Domain-Driven Generative Data Model for BigPetStore

### 2014-07-16 bdcloud_paper/latex/paper.tex
Updated figures.  Add equations.  Edited model description

Store records contain a unique identifier and a location in the form of a 5-digit zip code.  The locations of the stores are modeled by a PMF that gives the probability that a store’s location is the given zip code. We’ve written the PMF to give to zip codes in high-population, high-income areas the highest probabilities. The PMF is composed of two individual PMFs. One PMF determines the probability of each zip code as the population of that zip code over the total population:

@@

@@

where Z is the normalization factor.  Given the small size (~35,000 zip codes) of the data set, Z can be found directly by iterating over all of the zip codes and summing the scores.

@@

\begin{figure*}[!t]
  \centering
  \includegraphics[width=5.0in]{figures/transaction_modelseps.eps}
  \caption{Flowchart of Transaction Time \& Purchases Simulations}
  \label{fig:fig_sim}
\end{figure*}

@@

A transaction model is created for each customer and integrated over the simulation's timeframe to generate transactions.  Transactions are triggered when a customer runs out of items such as dog food as well as a random process describing the frequency of "impromptu" shopping trips.

Transaction times are modeled using a Monte Carlo method (Figure 1a).  The usage over time of each category is simulated, from which transaction times are found for each item category. The earliest transaction time is taken as the proposed transaction time.  The probability of the transaction time is calculated using a PDF.  If rejected, new transaction times are proposed.  Otherwise, the purchased items are modeled using a separate process (discussed below). After the items are chosen, the process begins again with the simulation of the item category usages.

@@

where $a_i$ is the amount of item category $i$ ``stuff'' remaining at time $t$. The variable $u_i$ gives the average usage rate and $\sigma^2_i$ gives the variance of the usage rate for item category $i$. From the usage simulations, the exhaustion time for each exhaustible item category is found.

The transaction times are sampled from a collection of Poisson processes, one for each item category as well as one for impromptu shopping trips, modeling the arrival time relative to the exhaustion time.  In other words, the time at which the customer visits the store is found by sampling arrival times from an exponential distribution (parameterized separately for each item category) and subtracting the arrival times from the exhaustion times.  The earliest arrival time is taken as the proposed transaction time.

The acceptance probability of the proposed transaction time is modeled using a PDF that incorporates factors such as store hours.  The PDF could be extended to incorporate additional factors including time-dependent events such as sales, weather, and the amount of money the customer has available at the time.

Until the transaction time is accepted, the process repeatedly samples new transaction times from the Poisson processes. If the transaction time is accepted, a separate model (discussed below) is used to choose the items purchased.  Afterwards, the process starts over with the simulation of the items' usages. The process is repeated until a transaction later than the end time of the simulation is generated.

\subsubsection{Purchased Items}

A customer’s purchases in each transaction are modeled using a Hidden Markov Model (HMM) for the item categories layered on top of separate Markov Models for each item category (Figure 1b).  The current state is the combination of the customer’s current inventory and purchasing preferences. The observables are the item categories and a "stop" state. The observables are assigned probabilities based on which items the user is running out of (based on the usage simulations) and the average number of items purchased per transaction. The item category HMM could be extended to consider additional factors such as the amount of money spent so far in the transaction and the customer's spending habits.

@@

>>>>

Store records contain a unique identifier and a location in the form of a 5-digit zip code.  The locations of the stores are modeled by a PMF that gives the probability that a store's location is the given zip code. We've written the PMF to give to zip codes in high-population, high-income areas the highest probabilities. The PMF is composed of two individual PMFs. One PMF determines the probability of each zip code as the population of that zip code over the total population:

\begin{equation}
p(\text{location}=z | \text{population}(z)) = \frac{\text{population}(z)}{\sum_{i} \text{population}(i)}
\end{equation}

@@

\begin{equation}
p(\text{location}=z | \text{income}(z)) = s ^ {\big( \frac{\text{income} - \min_i{\textrm{income}(i)}} {\max_i{\textrm{income}(i)} - \min_i{\textrm{income}(i)}} - 1 \big)}
\end{equation}

@@

\begin{align}
p(\text{location}=z) = &Z^-1 p(\text{location}=z | \text{population}(z)) \nonumber \\
&p(\text{location}=z | \text{income}(z))
\end{align}

where $Z$ is the normalization factor.  Given the small size ($\approx$30,000 zip codes) of the data set, $Z$ can be found directly by iterating over all of the zip codes and summing the scores.

@@

\begin{figure}[!t]
  \centering
  \includegraphics[width=3.75in]{figures/transaction_simulation.eps}
  \caption{Flowchart of Transaction Monte Carlo Simulation}
  \label{fig:trans_sim}
\end{figure}

@@

A transaction model is created for each customer and integrated over the simulation's timeframe to generate transactions.  Transactions are triggered when a customer runs out of items such as dog food as well as a random process describing the frequency of ``impromptu'' shopping trips.

Transaction times are simulated using a Monte Carlo method (Figure~\ref{fig:trans_sim}).  The usage over time of each category is simulated, from which transaction times are found for each item category. The earliest transaction time is taken as the proposed transaction time.  The probability of the transaction time is calculated using a PDF.  If rejected, new transaction times are proposed.  Otherwise, the purchased items are modeled using a separate process (discussed below). After the items are chosen, the process begins again with the simulation of the item category usages.

@@

\begin{equation}
\frac{da_i}{dt} = -\mu_i + \sigma_i\, dW_i(t)
\end{equation}

where $a_i$ is the amount of item category $i$ ``stuff'' remaining at time $t$. The variable $u_i$ gives the average usage rate and $\sigma^2_i$ gives the variance of the usage rate for item category $i$. The exhaustion time $ T_{E,i}$ for each item category is found from the usage simulations.

The proposed transaction time is found from the exhaustion times.  For each item cateogry, the offset time $T_{O, i}$ between when a customer would visit the store and the exhaustion time is sampled from an Exponential distribution parameterized separately for each customer $c$. A proposed transaction time $T_{P, i}$ for each item category is found by subtracting the offset time $T_{O, i}$ from the exhaustion time $T_{E, i}$.  The earlist proposed transaction time is taken as the overall proposed transaction time $T_T$.

\begin{align}
&\beta_c \sim \text{U}(a, b) \\
&T_{O, i} \sim \text{Exp}(\beta^{-1}_c) \\
&T_{P, i} = T_{E,i} - T_{P, i} \\
&T_T = \min_i \, \{  T_{P, i}\}
\end{align}

The acceptance probability $p(t=T_T)$ of the proposed transaction time $T_T$ is modeled using a PDF that incorporates factors such as store hours.  The PDF could be extended to incorporate additional factors including time-dependent events such as sales, weather, and the amount of money the customer has available at the time.

\begin{equation}
p(t=T_T) = p(t=T_T | \text{store hours})
\end{equation}

The proposed transaction time is accepted if $p(t=T_T) < r$, where $r \sim \text{U}(0, 1)$. Until the proposed transaction time is accepted, a new proposed transaction time is generated by sampling new offset times. If the proposed transaction time is accepted, the shopping cart of items is generated through a separate simulation discussed below. Transactions are generated until the accepted transaction time is later than the end time given as a simulation parameter.

\begin{figure*}[!t]
  \centering
  \includegraphics[width=6in]{figures/shopping_cart_simulation.eps}
  \caption{Shopping Cart Hidden Markov Model}
  \label{fig:shopping_cart_sim}
\end{figure*}

\subsubsection{Shopping Cart Items}

A customer’s purchases in each transaction are modeled using a layered Hidden Markov Model (HMM) (Figure~\ref{fig:shopping_cart_sim}).  The HMM has three states: the start state, the purchase states, and the stop state. There are an infinite number of purchase states, identified by the exhaustion times of the customer's item categories, the transaction time, and the number of items purchased in the current transaction. The observables of the purchase states correspond to the item categories and are Markov models for choosing the item to be purchased from the category. The observables are assigned probabilities based on the amounts of time (based on the usage simulations) until the customer uses the remaining amount of each item category.

Markov Models are used to model the customer's purchasing behavior for each item category.  Each state corresponds to an item in that category. The transition probabilities are defined so-as to model the customer's buying preferences.

\begin{equation}
Pr(X_{n+1}=x|X_n) = \dots
\end{equation}

The item category HMM could be extended to consider additional factors such as the amount of money spent so far in the transaction and the customer's spending habits.

@@

\subsubsection{Other Fields}
